---
title: "exercises01"
author: "Rodolfo León"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

1.  How many rows are in penguins? How many columns?

```{r}
dim(penguins)
```

2.  What does the bill_depth_mm variable in the penguins data frame describe? Read the help for ?penguins to find out.

The bill_depth_mma is a number denoting bill depth (millimeters)

3.  Make a scatterplot of bill_depth_mm vs. bill_length_mm. That is, make a scatterplot with bill_depth_mm on the y-axis and bill_length_mm on the x-axis. Describe the relationship between these two variables.

```{r}
ggplot(data = penguins, 
       aes(x = bill_length_mm, y = bill_depth_mm)) +
       geom_point()

```

4.  What happens if you make a scatterplot of species vs. bill_depth_mm? What might be a better choice of geom?

```{r}
ggplot(data = penguins, 
       aes(x = bill_depth_mm, y = species)) +
       geom_point()
```

5.  Why does the following give an error and how would you fix it?}

ggplot(data = penguins) + geom_point()

Answer: It lacks the x and y coordinates

```{r}
#ggplot(data = penguins,
#       ) + 
#  geom_point()
```

6.  What does the `na.rm` argument do in `geom_point()`? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to `TRUE`.

```{r}
ggplot(
  data = penguins, 
  aes(x = bill_depth_mm, y = bill_length_mm)
) + 
  geom_point(na.rm = TRUE)

# It removes the missing values without warning
```

7.  Add the following caption to the plot you made in the previous exercise: “Data come from the palmerpenguins package.” Hint: Take a look at the documentation for labs().

```{r}
ggplot(
  data = penguins, 
  aes(x = bill_depth_mm, y = bill_length_mm)
) + 
  geom_point(na.rm = TRUE) +
  labs(caption = "(Data come from the palmerpenguins package.)")
```

8.  Recreate the following visualization. What aesthetic should bill_depth_mm be mapped to? And should it be mapped at the global level or at the geom level?

```{r}
ggplot(
  data = penguins,
  aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = bill_depth_mm)) + 
  geom_smooth()

```

9. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)
  
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)

```
  
