---
title: "Row Exercises"
author: "Rodolfo León"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

```{r}
#| echo: false

```

1.  In a single pipeline for each condition, find all flights that meet the condition:

    -   Had an arrival delay of two or more hours

    -   Flew to Houston (`IAH` or `HOU`)

    -   Were operated by United, American, or Delta

    -   Departed in summer (July, August, and September)

    -   Arrived more than two hours late, but didn’t leave late

    -   Were delayed by at least an hour, but made up over 30 minutes in flight

        ```{r}

        ```

2.  Sort `flights` to find the flights with longest departure delays. Find the flights that left earliest in the morning.

    ```{r}

    ```

3.  Sort `flights` to find the fastest flights. (Hint: Try including a math calculation inside of your function.)

    ```{r}

    ```

4.  Was there a flight on every day of 2013?

    ```{r}

    ```

5.  Which flights traveled the farthest distance? Which traveled the least distance?

    ```{r}

    ```

6.  Does it matter what order you used [`filter()`](https://dplyr.tidyverse.org/reference/filter.html) and [`arrange()`](https://dplyr.tidyverse.org/reference/arrange.html) if you’re using both? Why/why not? Think about the results and how much work the functions would have to do.

    ```{r}

    ```
