---
title: "Quarto Template"
author: "Rodolfo León"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
#| echo: false
library(tidyverse)
library(ggthemes)
library(nycflights13)
```

1.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related?

2.  Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.

3.  What happens if you specify the name of the same variable multiple times in a [`select()`](https://dplyr.tidyverse.org/reference/select.html) call?

4.  What does the [`any_of()`](https://tidyselect.r-lib.org/reference/all_of.html) function do? Why might it be helpful in conjunction with this vector?

    ```         
    variables <- c("year", "month", "day", "dep_delay", "arr_delay")
    ```

5.  Does the result of running the following code surprise you? How do the select helpers deal with upper and lower case by default? How can you change that default?

    ```         
    flights |> select(contains("TIME"))
    ```

6.  Rename `air_time` to `air_time_min` to indicate units of measurement and move it to the beginning of the data frame.

7.  Why doesn’t the following work, and what does the error mean?

    ```         
    flights |>    select(tailnum) |>    arrange(arr_delay) #> Error in `arrange()`: #> ℹ In argument: `..1 = arr_delay`. #> Caused by error: #> ! object 'arr_delay' not found
    ```
