```{r}

```

---
title: "Groups"
author: "Rodolfo León"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
#| echo: false
library(tidyverse)
library(ggthemes)
library(nycflights13)
```

## 1. group_by()

```{r}
flights |> 
  group_by(month)
```

## summarize()

```{r}
flights |> 
  group_by(month) |> 
  summarize(avg_delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
flights |> 
  group_by(month) |> 
  summarize(
    avg_delay = mean(dep_delay, na.rm = TRUE), 
    n = n()                             # counts the number of rows in each group
  )
```

## The slice\_ functions()

---
There are five handy functions that allow you extract specific rows within each group:

df |> slice_head(n = 1) takes the first row from each group.
df |> slice_tail(n = 1) takes the last row in each group.
df |> slice_min(x, n = 1) takes the row with the smallest value of column x.
df |> slice_max(x, n = 1) takes the row with the largest value of column x.
df |> slice_sample(n = 1) takes one random row.
---

```{r}
flights |> 
  group_by(dest) |>              # Groups by destination
  slice_max(arr_delay, n = 1) |> # Extracts the row with the highest value 
  relocate(dest)
```

## Group by multiple variables

```{r}
daily <- flights |>  
  group_by(year, month, day)
daily
```

```{r}
daily_flights <- daily |> 
  summarize(n = n())

```

```{r}
daily_flights <- daily |> 
  summarize(
    n = n(), 
    .groups = "drop_last"
  )
# Explanation for these two in page 107 of the book
```

## Ungrouping

```{r}
daily |> 
  ungroup()
```

## Using .by

```{r}
flights |> 
  summarize(
    delay = mean(dep_delay, na.rm = TRUE), 
    n = n(),
    .by = c(origin, dest)
  )
```
