---
title: "Visualizing Distributions"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

```{r}
#| echo: false
```

The `echo: false` option disables the printing of code (only output is displayed).

### Bar Chart

```{r}
ggplot(penguins, aes(x = species)) +
geom_bar() 
```

```{r}
p <- ggplot(penguins, aes(x = fct_infreq(species)))
p + geom_bar() 
```

### Histogram

```{r}
# Numerical Variable
p <- ggplot(penguins, aes(x = body_mass_g))
p + geom_histogram(binwidth = 200)
```

```{r}
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 20)
```

```{r}
ggplot(penguins, aes(x = body_mass_g)) +
  geom_density()
```

### Exercises

1.  Make a bar plot of `species` of `penguins`, where you assign `species` to the `y` aesthetic. How is this plot different?

```{r}

```

2.  How are the following two plots different? Which aesthetic, `color` or `fill`, is more useful for changing the color of bars?

Answer: 'fill' is better for the color of the bars as it paints the whole bars, not just the borders, as 'color' does.

```{r}
 ggplot(penguins, aes(x = species)) +
          geom_bar(color = "red")
```

```{r}
 ggplot(penguins, aes(x = species)) +
         geom_bar(fill = "red")
```

3.  What does the `bins` argument in [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html) do?

    *The 'bins' argument sets the with, or amount of data represented in each bar.*

4.  Make a histogram of the `carat` variable in the `diamonds` dataset that is available when you load the tidyverse package. Experiment with different binwidths. What binwidth reveals the most interesting patterns?

    *The 0.1 binwidth is clear enough.*

    ```{r}
    ggplot(diamonds, aes(x = carat)) +
      geom_histogram(binwidth = 0.1) + labs(x = "0.1 carats")
    ```

```{r}
ggplot(diamonds, aes(x = carat)) +
  geom_histogram(binwidth = 0.01)+ labs(x = "0.01 carats")
```

```{r}
ggplot(diamonds, aes(x = carat)) +
  geom_histogram(binwidth = 0.05)+ labs(x = "0.05 carats")
```
