---
title: "Exercises 2"
author: "Rodolfo León"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

```{r}
#| echo: false

```

1.  The `mpg` data frame that is bundled with the ggplot2 package contains 234 observations collected by the US Environmental Protection Agency on 38 car models. Which variables in `mpg` are categorical? Which variables are numerical? (Hint: Type [`?mpg`](https://ggplot2.tidyverse.org/reference/mpg.html) to read the documentation for the dataset.) How can you see this information when you run `mpg`?

    *a) Categorical are manufacturer, model, trans, drv, fl and class;*

    *b) Numerical are displ, year, cyl cty, hwy*

    c\) Type head(mpg)

2.  Make a scatterplot of `hwy` vs. `displ` using the `mpg` data frame. Next, map a third, numerical variable to `color`, then `size`, then both `color` and `size`, then `shape`. How do these aesthetics behave differently for categorical vs. numerical variables?

    *The difference is a numerical variable doesn’t work with shape aesthetic but a categorical variable does. Also, the color scale is different for numerical and categorical variables.*

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point()

```

```{r}
ggplot(
  mpg, 
  aes(x = hwy, y = displ, size = cty)
) + 
  geom_point()
```

```{r}
ggplot(
  mpg, 
  aes(x = hwy, y = displ, size = cty, color = cty)
) + 
  geom_point()
```

```{r}
ggplot(
  mpg, 
  aes(x = hwy, y = displ, size = cty, color = cty, shape = drv)
) + 
  geom_point()
```

3.  In the scatterplot of `hwy` vs. `displ`, what happens if you map a third variable to `linewidth`?

    *Nothing would happen since no line has been set.*

<!-- -->

4.  What happens if you map the same variable to multiple aesthetics?

    *It works put it doesn't provide any useful information.*

```{r}
ggplot(mpg, aes(x = hwy, y = hwy, color = hwy)) + 
  geom_point()
```

5.  Make a scatterplot of `bill_depth_mm` vs. `bill_length_mm` and color the points by `species`. What does adding coloring by species reveal about the relationship between these two variables? What about faceting by `species`?

    *Adelies tend to have a shorter bill length and but deeper than Gentoo. Chinstrap have the greater bill_length and a bill depth similar to Adelie. Gento have the shortest bill depth and a bill length slightly under Chinstrap.*

    ```{r}
    ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
      geom_point()

    ```

6.  Why does the following yield two separate legends? How would you fix it to combine the two legends?

    *The code provided in the exercise yields two separate legends because the legend for `color` is renamed to `"Species"` but the legend for shape is not, and is named `"species"` by default instead. To fix it, we would need to explicitly rename the shape legend as well. Another option is to just remove shape = species to eliminate the duplication.*

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(
        x = bill_length_mm, y = bill_depth_mm, 
        color = species, shape = species
      )
    ) +
      geom_point() +
      labs(color = "Species",
           shape = "Species")
    ```

7.  Create the two following stacked bar plots.

    Which question can you answer with the first one? *The proportion of a species inhabiting the island*

    Which question can you answer with the second one? Same

    ```{r}
    ggplot(penguins, aes(x = island, fill = species)) +
      geom_bar(position = "fill")
    ```

```{r}
ggplot(penguins, aes(x = species, fill = island)) +
  geom_bar(position = "fill")
```
