---
title: "Layers"
author: "Rodolfo León"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
#| echo: false
suppressPackageStartupMessages(library(tidyverse))
library(ggthemes)
```

## Aesthetic mappings

```{r}
# Left
ggplot(mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point()

# Right
ggplot(mpg, aes(x = displ, y = hwy, shape = class)) +
  geom_point()

```

```{r}

```

```{r}
# Left
ggplot(mpg, aes(x = displ, y = hwy, size = class)) +
  geom_point()
```

```{r}
# Right
ggplot(mpg, aes(x = displ, y = hwy, alpha = class)) +
  geom_point()
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(color = "blue")
```

## Exercises

1.  Create a scatterplot of `hwy` vs. `displ` where the points are pink filled in triangles.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(color = "pink", shape = 2)
```

2.  Why did the following code not result in a plot with blue points?

    [ggplot](https://ggplot2.tidyverse.org/reference/ggplot.html)(mpg) + [geom_point](https://ggplot2.tidyverse.org/reference/geom_point.html)([aes](https://ggplot2.tidyverse.org/reference/aes.html)(x = displ, y = hwy, color = "blue"))

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(color = "blue")
# Color must be set outside the aes mapping
```

3.  What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use [`?geom_point`](https://ggplot2.tidyverse.org/reference/geom_point.html))

    *The **stroke** argument is used to control the size of the edge/border of your point.* 

4.  What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`? Note, you’ll also need to specify x and y.

    *It creates a logical variable with values `TRUE` and `FALSE` for cars with displacement values below and above 5. In general, mapping an aesthetic to something other than a variable first evaluates that expression then maps the aesthetic to the outcome.*

```{r}
ggplot(mpg, aes(x = hwy, y = displ, color = displ < 5)) + 
  geom_point()
```

## Geometric Objects

```{r}
# Left
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()

# Right
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_smooth()
#> `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
```

```{r}
# Left
ggplot(mpg, aes(x = displ, y = hwy, shape = drv)) + 
  geom_smooth()

# Right
ggplot(mpg, aes(x = displ, y = hwy, linetype = drv)) + 
  geom_smooth()
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = drv)) + 
  geom_point() +
  geom_smooth(aes(linetype = drv))
```

```{r}
# Left
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth()

# Middle
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(group = drv))

# Right
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(color = drv), show.legend = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(aes(color = class)) + 
  geom_smooth()
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    color = "red"
  ) +
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    shape = "circle open", size = 3, color = "red"
  )
```

```{r}
# Left
ggplot(mpg, aes(x = hwy)) +
  geom_histogram(binwidth = 2)

# Middle
ggplot(mpg, aes(x = hwy)) +
  geom_density()

# Right
ggplot(mpg, aes(x = hwy)) +
  geom_boxplot()
```

```{r}
library(ggridges)

ggplot(mpg, aes(x = hwy, y = drv, fill = drv, color = drv)) +
  geom_density_ridges(alpha = 0.5, show.legend = FALSE)
#> Picking joint bandwidth of 1.28
```

## Exercises

```{r}

mpg |> ggplot(aes(x = displ, y = hwy)) +
  geom_point() + 
  geom_smooth()

mpg |> ggplot(aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(aes(group = drv))

```

```{r}
mpg |> ggplot(aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth(aes(group = drv))


mpg |> ggplot(aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth()

```

```{r}
mpg |> ggplot(aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth(aes(linetype = drv, ))


mpg |> ggplot(aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth()

```
