---
title: "Row Exercises"
author: "Rodolfo León"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(nycflights13)
```

```{r}
#| echo: false

```

1.  In a single pipeline for each condition, find all flights that meet the condition:

-   Had an arrival delay of two or more hours

```{r}
flights |> filter(arr_delay >= 120) |> arrange(desc(arr_delay))
```

-   Flew to Houston (`IAH` or `HOU`)

```{r}
flights |> filter(dest %in% c("IAH", "HOU"))
```

-   Were operated by United, American, or Delta

```{r}
flights |>  filter(carrier %in% c("UA", "AA", "DL"))
```

-   Departed in summer (July, August, and September)

```{r}
flights |> filter(month %in% c(7, 8, 9))
```

-   Arrived more than two hours late, but didn't leave late

```{r}
flights |> filter(arr_delay >= 120 & dep_delay <= 0) 
```

-   Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
flights |> 
  filter(arr_delay >= 120 & dep_delay <= 0)
```

2.  Sort `flights` to find the flights with longest departure delays. Find the flights that left earliest in the morning.

```{r}

```

3.  Sort `flights` to find the fastest flights. (Hint: Try including a math calculation inside of your function.)

```{r}

```

4.  Was there a flight on every day of 2013?

```{r}

```

5.  Which flights traveled the farthest distance? Which traveled the least distance?

```{r}

```

6.  Does it matter what order you used if you’re using both? Why/why not? Think about the results and how much work the functions would have to do.

```{r}

```
